Results

### 2024-05-23 manuscript.Rmd
- updated for revision of IPA

documentclass     : "apa6"
# classoption       : "man"
# output            : html_document
output            : papaja::apa6_pdf
# output            : papaja::apa6_word

@@

```{r setup, include=F}
devtools::install_github("tdienlin/td@v.0.0.2.5")  # uncomment to install
library(devtools)
library(english)
library(ggplot2)
library(kableExtra)
library(lavaan)
library(magrittr)
library(naniar)
library(papaja)
library(semPlot)
library(simsem)
library(td)
library(tidyverse)

@@

We wanted to find out _who_ needs privacy, and not so much _what causes_ the need for privacy.
Hence, to answer our research question, in a joint model combining all variables (including sociodemographic variables) we analyzed the variables' bivariate relations. 
To predict the need for privacy, we first used the six personality factors.
Afterward, we predicted privacy using the more specific facets. 
To get a first idea of the variables' potential causal relations, we also planned to run latent structural regression models.
However, because model fit was not acceptable, in exploratory analyses we investigated the potential effects in a multiple regression using the mean values of the observed scores.

@@

```{r tab-des, cache=F, warning=F, message=F, prompt=F}
apa_table(
  tab_des, 
  stub_indents = list(
    "Personality" = c(1:31),
    "Need for Privacy" = c(32:39)
    ),
  align = c("l", rep("r", 8)),
  caption = "Factorial validity of all measures.",
  font_size = "small",
  escape = FALSE
  )
```

@@

More agreeable participants showed a similar pattern.
They needed less psychological (`r get_res("Agreeableness", "Psychological")`), social privacy (`r get_res("Agreeableness", "Social")`), and physical privacy (`r get_res("Agreeableness", "Physical")`).
The facets showed virtually the same pattern.
Effect sizes were substantial, but on the whole smaller than those for extraversion.

@@

@@

In Table \@ref(tab:tab-dim), we report how the personality dimensions predicted need for privacy.
In Table \@ref(tab:tab-fac), we report how the personality facets predicted need for privacy.
In Table \@ref(tab:tab-ses), we report how sociodemographics predicted need for privacy.
Figure \@ref(fig:fig-cor) summarizes how all of the variables---dimensions, facets, and sociodemographics---predicted the need for privacy.

@@

In exploratory analyses we analyzed how personality facets might have potentially caused need for privacy.
We found that the need for psychological privacy was explained by two variables: extraversion (`r get_res("Extraversion", "Psychological", "tab_reg")`) and emotionality (`r get_res("Emotionality", "Psychological", "tab_reg")`). 
The need for social privacy was also potentially affected by extraversion and emotionality.
Being more extraverted substantially decreased the need for psychological privacy (`r get_res("Extraversion", "Social", "tab_reg")`), as did being more emotional (`r get_res("Emotionality", "Social", "tab_reg")`).
Physical privacy was determined by again extraversion, but also by agreeableness and conscientiousness.
Being more extraverted appeared to decrease the need for physical privacy (`r get_res("Extraversion", "Physical", "tab_reg")`); 
being more agreeable likewise decreased the need for physical privacy (`r get_res("Agreeableness", "Physical", "tab_reg")`);
however, being more conscientious increased the need for physical privacy (`r get_res("Conscientiousness", "Physical", "tab_reg")`).
The need for privacy from the government was affected by the two factors of openness and conservatism.
Being more open to new experiences potentially increased the need for privacy from the government (`r get_res("Openness", "Government", "tab_reg")`), as did being more politically conservative (`r get_res("Conservatism", "Government", "tab_reg")`).
The need for privacy from companies was affected by the openness to new experiences only. 
Being more open to new experiences potentially increased the need for privacy from companies (`r get_res("Openness", "Companies", "tab_reg")`).
Being extraverted and conscientious affected the need for informational privacy.
Whereas being more extraverted decreased the need for informational privacy (`r get_res("Extraversion", "Informational", "tab_reg")`), being more conscientious increased the need for informational privacy (`r get_res("Conscientiousness", "Informational", "tab_reg")`) in our data.
The need for anonymity was meaningfully affected only by extraversion.
More extraverted people need less anonymity (`r get_res("Extraversion", "Anonymity", "tab_reg")`).
Finally, the general need for privacy was affected by four variables.
Being extraverted again decreased the general need for privacy (`r get_res("Extraversion", "General", "tab_reg")`).
However, the general need for privacy was increased by being more conscientious (`r get_res("Conscientiousness", "General", "tab_reg")`), more conservative (`r get_res("Conservatism", "General", "tab_reg")`), and more open to experiences (`r get_res("Openness", "General", "tab_reg")`).

@@

In addition, we looked at relations between need for privacy and various sociodemographic variables.
Contrary to our expectations, older participants desired less privacy from others, both socially and physically. 
It is an open question as to whether this relation represents a developmental mechanism or a difference between cohorts.
Research suggests that older people have fewer social interactions than younger people [@ortiz-ospinaTimeUse2024], which could result in a lower need for social and physical privacy.
But it could also be a difference between cohorts. 
Younger people nowadays have fewer in-person social contacts than before, often attributed to increases in time spent online [@twengeLessInpersonSocial2019].
Hence, it could also be that younger generations prefer more solitude than the generations beforehand.

@@

Based on correlations, the results above analyzed the variables' relationships.
To analyze the potential impact of personality on need for privacy, in exploratory analyses we also ran several multiple regression analyses, in which we controlled for the respective personality variables and sociodemographics.
Results were comparable, in that extraversion turned out to be the major potential cause of need for privacy.
However, there were also some differences.
Most notably, results implied that being more conscientious increases the need for physical, informational, and general privacy.
Similarly, being more open to new experiences might increase the need for privacy from the government, from companies, and for privacy in general.
Finally, multiple regression results suggested that being conservative does not only increase the need for privacy from the government but also for privacy in general.
All other sociodemographic variables ceased to be significant.

Looking at the results more broadly, we make four general observations.
First, it makes sense to differentiate different levels of need for privacy.
Many personality traits showed meaningful relations with some dimension, yet no or even opposite relations with others.
To illustrate, whereas more conscientious people desired less psychological privacy and less privacy from other, when asked about privacy in general and privacy from companies they needed more. 
General need for privacy may rather represent a cognitive appraisal, whereas social or psychological privacy might be more experienced-based and psychological. 
In any case, our results argue for a more rather than less nuanced strategy for measuring privacy attitudes.

@@

This leads to our fourth general observation. 
Our main interest was to determine the personality factors predicting the need for privacy. 
What can we learn about a person given their need for privacy? 
We first analyzed this question in correlation analyses. 
But in additionally exploring potential causal effects in a multiple regression, a somewhat different picture emerged. 
The results suggest that several of the correlations we found might not be the result of a causal processes. 
At the same time, some causal effects might also be suppressed in correlation analyses.
More research better suited at uncovering causal relations is needed to address this question.

@@

Instead of deleting items or changing the factor structure, to avoid over-fitting we decided to maintain the measures’ original factor structure.
For this reason, instead of reporting the results from latent structural equation modelling, we reported the results of the correlations of the observed variables' means. 
Interested readers can find the results of the latent analyses in our online supplementary material.
The results are highly comparable, with the major difference that effect sizes in the latent models tended to be larger.
The results reported here are hence more conservative and likely underestimate the true effect sizes.
This underscores the need for further optimizing measures of personality and privacy. 
We hope this study is a catalyst for future work in this area.

>>>>

documentclass     : "apa6"
# classoption       : "man"
output            : html_document
# output            : papaja::apa6_pdf
# output            : papaja::apa6_word

@@

```{r setup, include=F}
# devtools::install_github("tdienlin/td@v.0.0.2.5")  # uncomment to install
library(devtools)
library(english)
library(ggplot2)
library(kableExtra)
library(lavaan)
library(magrittr)
library(naniar)
library(papaja)
library(semPlot)
library(simsem)
library(td)
library(tidyverse)

@@

We wanted to find out _who_ needs privacy, and not so much _what causes_ the need for privacy.
Hence, to answer our research question, in a joint model combining all variables (including sociodemographic variables) we analyzed the variables' bivariate relations. 
To predict the need for privacy, we first used the six personality factors.
Afterward, we predicted privacy using the more specific facets. 
To get a first idea of the variables' potential causal relations [@groszTabooExplicitCausal2020a], we also planned to run latent structural regression models.
However, because model fit was not acceptable, in exploratory analyses we investigated the potential effects in a multiple regression using the mean values of the observed scores.

@@

```{r tab-des, cache=F, warning=F, message=F, prompt=F}
apa_table(
  tab_des, 
  stub_indents = list(
    "Personality" = c(1:31),
    "Need for Privacy" = c(32:39)
    ),
  align = c("l", rep("r", 8)),
  caption = "Factorial validity of all measures.",
  font_size = "footnotesize",
  escape = FALSE,
  note = "REL: Reliability measured via McDonald's Omega; CFI = Comparative Fit Index; TLI = Tucker-Lewis Index; SRMR = Standardized Root Mean Square Residual; RMSEA = Root Mean Square Error of Approximation"
  )
```

@@

More agreeable participants showed a similar pattern.
They needed less psychological (`r get_res("Agreeableness", "Psychological")`), social (`r get_res("Agreeableness", "Social")`), and physical privacy (`r get_res("Agreeableness", "Physical")`).
The facets showed virtually the same pattern.
Effect sizes were substantial, but on the whole smaller than those for extraversion.

@@

In Table \@ref(tab:tab-dim), we report how the personality dimensions predicted need for privacy.
In Table \@ref(tab:tab-fac), we report how the personality facets predicted need for privacy.

In Table \@ref(tab:tab-ses), we report how sociodemographics predicted need for privacy.
Figure \@ref(fig:fig-cor) summarizes how all of the variables---dimensions, facets, and sociodemographics---predicted the need for privacy.

@@

@@

In exploratory analyses we analyzed how personality facets might have potentially caused need for privacy, using multiple regressions in which we controlled for all personality dimensions and sociodemographic variables. 
We found that the need for psychological privacy was explained by two variables: extraversion (`r get_res("Extraversion", "Psychological", "tab_reg")`) and emotionality (`r get_res("Emotionality", "Psychological", "tab_reg")`). 
The need for social privacy was also potentially affected by extraversion and emotionality.
Being more extraverted substantially decreased the need for psychological privacy (`r get_res("Extraversion", "Social", "tab_reg")`), as did being more emotional (`r get_res("Emotionality", "Social", "tab_reg")`).
Physical privacy was determined by again extraversion, but also by agreeableness and conscientiousness.
Being more extraverted appeared to decrease the need for physical privacy (`r get_res("Extraversion", "Physical", "tab_reg")`); 
being more agreeable likewise decreased the need for physical privacy (`r get_res("Agreeableness", "Physical", "tab_reg")`);
however, being more conscientious increased the need for physical privacy (`r get_res("Conscientiousness", "Physical", "tab_reg")`).
The need for privacy from the government was affected by the two factors of openness and conservatism.
Being more open to new experiences potentially increased the need for privacy from the government (`r get_res("Openness", "Government", "tab_reg")`), as did being more politically conservative (`r get_res("Conservatism", "Government", "tab_reg")`).
The need for privacy from companies was affected by the openness to new experiences only. 
Being more open to new experiences potentially increased the need for privacy from companies (`r get_res("Openness", "Companies", "tab_reg")`).
Being extraverted and conscientious affected the need for informational privacy.
Whereas being more extraverted decreased the need for informational privacy (`r get_res("Extraversion", "Informational", "tab_reg")`), being more conscientious increased the need for informational privacy (`r get_res("Conscientiousness", "Informational", "tab_reg")`) in our data.
The need for anonymity was meaningfully affected only by extraversion.
More extraverted people need less anonymity (`r get_res("Extraversion", "Anonymity", "tab_reg")`).
Finally, the general need for privacy was affected by four variables.
Being extraverted again decreased the general need for privacy (`r get_res("Extraversion", "General", "tab_reg")`).
However, the general need for privacy was increased by being more conscientious (`r get_res("Conscientiousness", "General", "tab_reg")`), more conservative (`r get_res("Conservatism", "General", "tab_reg")`), and more open to experiences (`r get_res("Openness", "General", "tab_reg")`).

@@

In addition, we looked at relations between need for privacy and various sociodemographic variables.
Contrary to our expectations, older participants desired less privacy from others, both socially and physically. 
This was surprising, for example given that older people report increased online privacy concern [@kezerAgeDifferencesPrivacy2016].
It is an open question as to whether this relation represents a developmental mechanism or a difference between cohorts.
Research suggests that older people have fewer social interactions than younger people [@ortiz-ospinaTimeUse2024], which could result in a lower need for social and physical privacy.
But it could also be a difference between cohorts. 
Younger people nowadays have fewer in-person social contacts than before, often attributed to increases in time spent online [@twengeLessInpersonSocial2019].
Hence, it could also be that younger generations prefer more solitude than the generations beforehand.

@@

The results above are based on correlations and analyzed the variables' relationships.
To analyze the potential impact of personality on need for privacy, in exploratory analyses we also ran several multiple regression analyses.
Here, we estimated the relations between each personality and need for privacy dimension while controlling for all other personality dimensions and sociodemographics.
Results were comparable, in that extraversion turned out to be the major potential cause of need for privacy.
However, there were also some differences.
Most notably, results implied that being more conscientious increases the need for physical, informational, and general privacy.
Similarly, being more open to new experiences might increase the need for privacy from the government, from companies, and for privacy in general.
Finally, multiple regression results suggested that being conservative does not only increase the need for privacy from the government but also for privacy in general.
All other sociodemographic variables ceased to be significant.

Looking at the results more broadly, we make five general observations.
First, it makes sense to differentiate different levels of need for privacy.
Many personality traits showed meaningful relations with some dimension, yet no or even opposite relations with others.
To illustrate, whereas more conscientious people desired less psychological privacy and less privacy from other, when asked about privacy in general and privacy from companies they needed more. 
General need for privacy may rather represent a cognitive appraisal, whereas social or psychological privacy might be more experienced-based and psychological. 
In any case, our results argue for a more rather than less nuanced strategy for measuring privacy attitudes.

@@

This leads to our fourth general observation. 
Our main interest was to determine the personality factors predicting the need for privacy. 
What can we learn about a person given their need for privacy? 
We first analyzed this question in correlation analyses. 
However, when we further explored potential causal effects using multiple regression analysis [@groszTabooExplicitCausal2020a], a somewhat different picture emerged. 
Several of the bivariate relationships we initially observed disappeared. 
These results suggest that some of the correlations we found might not be due to a direct causal process but could instead be explained by shared variance with a third confounding factor. 
Additionally, some causal effects that were not apparent in the correlation analyses became significant in the multiple regressions. 
Specifically, conservatism became more relevant when we included additional control variables, suggesting that an increase in conservatism leads to a greater desire for anonymity and privacy in general.

Finally, our results have significant implications for privacy theory, suggesting new avenues for further research. 
Given the close relationship between personality and privacy, it is worthwhile to examine how personality traits interact with specific privacy models and theories. 
For instance, personality could influence the privacy calculus model [@kezerGettingPrivacyCalculus2022]. 
Extraverted individuals, who tend to view others as resources, might perceive greater benefits in sharing information online and therefore be more open in their sharing behaviors. 
Conversely, anxious individuals, who are more likely to see others as threats, might have heightened concerns about privacy, leading to a reduced willingness to share information online.

@@

Instead of deleting items or changing the factor structure, to avoid over-fitting we decided to maintain the measures’ original factor structure.
For this reason, instead of reporting the results from latent structural equation modelling, we reported the results of the correlations of the observed variables' means. 
Interested readers can find the results of the latent analyses in our online supplementary material.
The results are highly comparable, with the major difference that effect sizes in the latent models tended to be larger.
The results reported here are hence more conservative and likely underestimate the true effect sizes.
This underscores the need for further optimizing measures of personality and privacy.

In our exploratory analyses, we aimed to investigate the potential causal effects of personality on the need for privacy by controlling for potential confounders, including other personality dimensions and sociodemographic variables. 
However, it is important to note that this is only a preliminary approach, as there are likely additional variables that could further explain the relationships. 
Future research should explore the potential causal relationships in a more systematic and comprehensive manner.

Our findings are specific to the U.S. context. 
Previous research indicates that although privacy is a universal concept, attitudes and practices vary significantly across different cultures and countries [@altmanPrivacyRegulationCulturally1977]. 
It is possible that the relationship between personality and the need for privacy manifests differently in various settings. 
We hope this study will serve as a catalyst for future research in this area.

### 2024-05-23 results.RmD
- updated for revision of IPA

fig_cor <- ggplot(
  tab_cor, aes(x = Estimate, y = Predictor, color = relevant)
  ) + 
  coord_cartesian(xlim = c(-.7, .3)) +
  geom_vline(xintercept = 0, lwd = .5, colour = "darkgrey") + 
  geom_vline(xintercept = -.1, lwd = .5, colour = "darkgrey", linetype = "dashed") + 
  geom_vline(xintercept = .1, lwd = .5, colour = "darkgrey", linetype = "dashed") + 
  geom_errorbarh(
    aes(xmin = CI_low, xmax = CI_high),
    lwd = .5, height = .4, position = position_dodge(-.7)
    ) +
  geom_point(size = 1, position = position_dodge(-.7)) +
#   geom_pointrange(aes(xmin = CI_low, xmax = CI_high), size = .2) + 
  scale_x_continuous(
    breaks = c(-.6, -.4, -.2, 0, .2, .4),
    labels = c("-.6", "-.4", "-.2", "0", ".2", ".4")
    ) + 
  scale_y_discrete(limits=rev) +
  labs(
    title = "Predicting the need for privacy",
    x = "Correlation coefficient"
    ) +
  # theme_minimal() +
  theme(
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = .5),
    text = element_text(size = 14),
    axis.text.y = element_text(hjust = 1),
    axis.text.x = element_text(size = 10),
    legend.position = "none"
    ) + 
  scale_color_manual(values = c("grey", "black")) +
  facet_grid(category ~ `Need for privacy`, scales = "free_y")

@@

fig_reg <- ggplot(
  tab_reg, aes(x = Estimate, y = Predictor, color = relevant)
  ) + 
  coord_cartesian(xlim = c(-.8, .3)) +
  geom_vline(xintercept = 0, lwd = .5, colour = "darkgrey") + 
  geom_vline(xintercept = -.1, lwd = .5, colour = "darkgrey", linetype = "dashed") + 
  geom_vline(xintercept = .1, lwd = .5, colour = "darkgrey", linetype = "dashed") + 
  geom_errorbarh(
    aes(xmin = CI_low, xmax = CI_high),
    lwd = .5, height = .4, position = position_dodge(-.7)
    ) +
  geom_point(size = 1, position = position_dodge(-.7)) +
  scale_x_continuous(
    breaks = c(-.8, -.6, -.4, -.2, 0, .2, .4),
    labels = c("-.8", "-.6", "-.4", "-.2", "0", ".2", ".4")
    ) + 
  scale_y_discrete(limits=rev) +
  scale_color_manual(values = c("grey", "black")) +
  labs(
    title = "Explaining the need for privacy",
    x = "Regression coefficient"
    ) +
  # theme_minimal() +
  theme(
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = .5),
    text = element_text(size = 14),
    axis.text.y = element_text(hjust = 1),
    axis.text.x = element_text(size = 10),
    legend.position = "none"
    ) + 
  facet_grid(category ~ `Need for privacy`, scales = "free_y")

>>>>

fig_cor <- ggplot(
  tab_cor, aes(x = Estimate, y = Predictor, color = relevant)
  ) + 
  coord_cartesian(xlim = c(-.8, .4)) +
  geom_vline(xintercept = 0, lwd = .5, colour = "darkgrey") + 
  geom_vline(xintercept = -.1, lwd = .5, colour = "darkgrey", linetype = "dashed") + 
  geom_vline(xintercept = .1, lwd = .5, colour = "darkgrey", linetype = "dashed") + 
  geom_errorbarh(
    aes(xmin = CI_low, xmax = CI_high),
    lwd = .5, height = .4, position = position_dodge(-.7)
    ) +
  geom_point(size = 1, position = position_dodge(-.7)) +
#   geom_pointrange(aes(xmin = CI_low, xmax = CI_high), size = .2) + 
  scale_x_continuous(
    breaks = c(-.8, -.6, -.4, -.2, 0, .2, .4),
    labels = c("-.8", "-.6", "-.4", "-.2", "0", ".2", ".4")
    ) + 
  scale_y_discrete(limits=rev) +
  labs(
    title = "Predicting the need for privacy",
    x = "Correlation coefficient"
    ) +
  # theme_minimal() +
  theme(
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = .5),
    text = element_text(size = 14),
    axis.text.y = element_text(hjust = 1),
    axis.text.x = element_text(size = 8),
    legend.position = "none"
    ) + 
  scale_color_manual(values = c("grey", "black")) +
  facet_grid(category ~ `Need for privacy`, scales = "free_y")

@@

fig_reg <- ggplot(
  tab_reg, aes(x = Estimate, y = Predictor, color = relevant)
  ) + 
  coord_cartesian(xlim = c(-.8, .3)) +
  geom_vline(xintercept = 0, lwd = .5, colour = "darkgrey") + 
  geom_vline(xintercept = -.1, lwd = .5, colour = "darkgrey", linetype = "dashed") + 
  geom_vline(xintercept = .1, lwd = .5, colour = "darkgrey", linetype = "dashed") + 
  geom_errorbarh(
    aes(xmin = CI_low, xmax = CI_high),
    lwd = .5, height = .4, position = position_dodge(-.7)
    ) +
  geom_point(size = 1, position = position_dodge(-.7)) +
  scale_x_continuous(
    breaks = c(-.8, -.6, -.4, -.2, 0, .2),
    labels = c("-.8", "-.6", "-.4", "-.2", "0", ".2")
    ) + 
  scale_y_discrete(limits=rev) +
  scale_color_manual(values = c("grey", "black")) +
  labs(
    title = "Explaining the need for privacy",
    x = "Regression coefficient"
    ) +
  # theme_minimal() +
  theme(
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = .5),
    text = element_text(size = 14),
    axis.text.y = element_text(hjust = 1),
    axis.text.x = element_text(size = 8),
    legend.position = "none"
    ) + 
  facet_grid(category ~ `Need for privacy`, scales = "free_y")
